Why do we Eval?\\[2mm]\Large A large-scale study of Eval usage in R

### 2020-11-26 inst/paper/main.tex

Most dynamic languages provide their users with a facility to
transform unstructured text into executable code, and to evaluate that
code. In this paper we refer to this reflective facility of a language
as \eval bowing to its origins in LISP in 1956~\cite{lisp}. \Eval has
been much maligned over the years. In computing lore, it is as close
to a boogeyman as it gets. Yet, for McCarthy, \eval was simply the way
to write down the definition of LISP, he was surprised that someone
coded it up and offered it end users.  Since then, reflective
facilities have been used to parameterize programs over code patterns
that can be provided after the program is written, when it
executes. The presence of such a feature in a language is a hallmark
of dynamism as it is a form of delayed binding.

@@

\vspace{2mm}\noindent\emph{Our contributions.} 
TODO

Our methodology for this work bows to the fact that static analysis of R is
difficult, instead we focus on dynamic analysis. To observe \eval we have
built a two-level monitoring infrastructure: we are able to monitor R
programs by instrumentation -- this gives us access to many user-visible
properties of R programs -- and we also can monitor the inner working of the
R interpreter -- this allows us to capture details of execution that are not
exposed at the source level.

Our corpus has been constructed to reflect the different level of
sophistication in the R community.

Thus, our
corpus distinguished between \emph{core packages} (there are
\CorpusCorePackages such packages that are distributed together with the R
virtual machine), \emph{CRAN packages} (we selected \CorpusPackages curated
packages that pass stringent community quality checks and are equipped with
tests and sample data), \emph{Kaggle scripts} (we used \CorpusFinishedKaggle
end-user written programs that performs a particular data analysis
task). There are reasons to believe that \eval usage patterns may differ
between these datasets. The core R code is extremely stable, it has only
been maintained for the last 20 years with very few far reaching
changes. The libraries represent a more lively ecosystem with new libraries
added each day.  Finally, end user code is often one-use scripts thrown
together and sometimes never revisited.

To understand the behavior of \eval we perform a dynamic analysis and manual
code inspection. The results presented here summarize our findings.

>>>>

Most dynamic languages provide their users with a facility to
transform unstructured text into executable code, and to evaluate that
code. In this paper we refer to this reflective facility of a language
as \eval bowing to its origins in LISP in 1956~\cite{lisp}. \Eval has
been much maligned over the years. In computing lore, it is as close
to a boogeyman as it gets. Yet, for McCarthy, \eval was simply the way
to write down the definition of LISP, he was surprised that someone
coded it up and offered it end users.  Since then, reflective
facilities have been used to parameterize programs over code patterns
that can be provided after the program is written.  The presence of
such a feature in a language is a hallmark of dynamism as it is a form
of delayed binding as what any particular call to \eval will do is
only know at program run-time when that particular call site is
evaluated.

@@

\vspace{2mm}\noindent\emph{The What and How.}  One significant benefit
of choosing R is that every package in the CRAN repository is curated
and comes with examples of typical usage. This gives us a large code
base that we can analyze dynamically.  To observe \eval we built a
two-level monitoring infrastructure: we can monitor R programs by
instrumentation --this gives us access to many user-visible properties
of R programs-- but we can also monitor the inner working of the R
interpreter --this allows us to capture details of execution that are
not exposed at the source level. Dynamic analysis is of course limited
in that it can only observe behaviors that are triggered by the
particular inputs passed to the program. Luckily, CRAN libraries come
with multiple input datasets and use-cases.

The choice of corpus is crucial for analyses like ours.  Our corpus
has been constructed to reflect the levels of sophistication of the R
community.  We distinguish between \emph{core packages} (there are
\CorpusCorePackages such packages that are distributed together with
the R virtual machine), \emph{CRAN packages} (\CorpusPackages curated
packages that pass stringent quality checks and are equipped with
tests and sample data), \emph{Kaggle scripts} (\CorpusFinishedKaggle
end-user written programs that performs a particular data analysis
task). It is reasonable to expect that \eval usage differ between
these datasets. The core R code is extremely stable, maintained for
the last 20 years with very few far reaching changes. The libraries
represent a more lively ecosystem with new libraries added each day.
Finally, end user code is often one-use scripts thrown together and
sometimes never revisited.
